---
title: "Final Report"
author: "Shujun Zhang"
date: "June 15, 2019"
output:
  pdf_document: default
  html_document:
    css: faded.css
---
## Data

The `gapminder` data set contains a collection of 6 variables measured between 1800 and 2015 on the world.It's describing life expentency depending on factors like GDP, Region, population etc. 

Some of the variables in the `gapminder` data set are:

* **life** - life expectancy
* **income** - gdp per capita
* **year** - the range from 1800 to 2015
* **county** - countries in the world
* **region** - 6 regions includes all contries
* **population** - census data collected about every 10 years

## Questions to Answer
1. Is there a big difference in terms of GPD per region? 
2. What's a potential relationship between a country’s GDP (income) and life expectancy?
3. For example in region America, how does it look like regarding to Life Expectancy Per Region in recent years?
 
## Data narrative summary
```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
rm(list = ls())
myData <- read.csv("D:\\ANLY_506\\final_project\\gapminder.csv", header=TRUE)
head(myData)
tail(myData) 
names(myData)
summary(myData)
nObs = nrow(myData)
nVars = ncol(myData)
year2015 <- filter(myData, Year == 2015)
```

1. There are **`r nObs`** observations in the data.
2. Thee are **`r nVars`** variables in the data.
3. Type of variables: "Country" is **`r class(myData$Country)`**, "Year" is **`r class(myData$Year)`**, "life" is **`r class(myData$life)`**, "population" is **`r class(myData$population)`**, "income" is **`r class(myData$income)`** and "region" is **`r class(myData$region)`**.
4. How disperse is the data: Range of "Year" is **`r range(myData$Year)`**. Range of "life" is **`r range(myData$life)`**. Range of "income" is **`r range(myData$income, na.rm = TRUE)`**.
5. Data wrangling: The avaerage life expectancy in year 2015 is **`r mean(year2015$life, na.rm = TRUE)`**.
6. Preprocessing steps: Bascially what I did was filling in the missing population data by the most recent non-missing values using 'fill' method. I also removed the rows with empty income values, and convert the type of population from factor to numeric type for later processing. Please check the in-line code for details.

```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
gm = filter(myData, !is.na(myData$income))
class(gm$population)
gm$population = as.numeric(as.character(gm$population))
class(gm$population)
gm <- gm %>% fill(population)
```

## Exploratory Plots
```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
ggplot(gm, aes(region, income, color = region)) + geom_boxplot() + labs(title="GDP Per Region", y = "GDP", x = "Region")
```

The above **Fig. 1** is a boxplot which shows the total GPD (income) per Region. 


```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
ggplot(data = gm, mapping = aes(x = income, y = life, color = region)) +
  geom_point(alpha = 0.5) +
  scale_x_log10()+
  scale_color_manual(values = c("red","yellow", "orange","green","blue","purple")) + 
  labs(title="The relation between GDP and Life Expenctancy", y = "Life Expectancy", x = "GDP")
```

The above **Fig. 2** is a scatter plot which shows the relation between GDP and Life Expenctancy. 

```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
gm2010 <- filter(gm, Year >= 2010)
ggplot(gm2010, aes(region, life, color = region)) + geom_boxplot() + labs(title="Life Expectancy Per Region since 2010", y = "Life Expectancy", x = "Region")
```

The above **Fig. 3** is a boxplot which shows the Life Expectancy Per Region in the recent years. 

## Clustering Analysis

**Note:** I only consider data in region America to save running time.

```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
library("factoextra")
# prepare data
gm_america <- filter(gm, region == "America")
gm_life_income_america <- select(gm_america, income, life)
# Silhouette Method
fviz_nbclust(gm_life_income_america, kmeans, method = "silhouette",nstart=20) # cluster 2
# Elbow Method
fviz_nbclust(gm_life_income_america, kmeans, method = "wss",nstart=20) # cluster 3
```

The **Fig. 4** (above two figures) is for finding the number of clusters using Silhouette Method and Elbow Method.

```{r, echo = FALSE, message=FALSE, results="hide", warning=FALSE}
# try 2 clusters and 3 clusters
k2 <- kmeans(gm_life_income_america, centers = 2)
k3 <- kmeans(gm_life_income_america, centers = 3)
# compare
fviz_cluster(k2, geom = "point", data = gm_life_income_america) + ggtitle("cluster = 2") + 
  xlab("GDP") + ylab("Life Expectancy")
fviz_cluster(k3, geom = "point",  data = gm_life_income_america) + ggtitle("cluster = 3") +
  xlab("GDP") + ylab("Life Expectancy")
```

The **Fig. 5** (above two figures) is the visualization of kmeans of 2 clusters and 3 clusters .

## Answers:
1. From **Fig. 1**, we could see differences between regions, especially for region Africa.
2. From **Fig. 2**, there's a positive relationship between a country’s GDP (income) and life expectancy.
3. From **Fig. 5**, lower income (GDP) will lead lower life expectancy, but when the income comes to a certain high level, the life expectancy won't increase too much.

## References:
1. Phillips, N. D. (2016). Yarrr! The pirate’s guide to R.
2. Peng, R. D. & Matsui, E. (2018). The Art of Data Science: A Guide for Anyone Who Works with Data. Skybrude Consulting, LLC. 
3. Grolemund, G., & Wichham, H. (2018). R for Data Science. 